Discrete Quantum Theories and Computing

### 2019-01-03 dissertation.tex
dissertation.lyx "Chapter 1 Introduction" Updating...

\centering{}\centering{}%
\begin{tabularx}{\linewidth}{ccYY}
\toprule 
\addlinespace
Physical quantities & Discrete & Discrete with extreme small unit & Continuous, or no evidence to support they are discrete\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational devices & Digital computer & Analog computer & Analog computer\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Mathematical representation & Integers & Real numbers & Real numbers\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational model  & Turing machine & BCSS machine & BCSS machine\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
How the models realize? & Reliably & 
Not reliably, but after we have a better technology to manipulate the minimum units, we can predict the physical phenomenon and their computational power by a better discrete model. & 
Not reliably, because precision can never be high enough, and the difference between their computational power might not be able compensated by error analysis techniques. \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Inspired quantum models &  & Quantum theories and computing over finite fields & Quantum interval-valued probability measures (QIVPM)\tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The situations for classical quantities and the computational power
above them are summarized in Table~\ref{tab:Comparison}, and the
two possibilities dealing with real-valued models inspires us two
types of quantum models. Similar to classical analog computers, the
quantum circuit model, the most widely used model of quantum computing,
manipulates quantities which are not discrete apparently: the probability
amplitudes of quantum states. To address the similar inconsistency,
we first review the ``conventional quantum theory'' (CQT)\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.} in Chapter~\ref{chap:Conventional-Quantum-Theory}, where the probability
amplitudes are in the field of complex numbers. If the probability
amplitudes are assumed to be ultimately discrete, since their field
structure is essential to construct a quantum theory, we might assume
they are elements in discrete finite fields and build the quantum
theories and computing upon these finite fields in Chapter~\ref{chap:Unrestricted-Finite-Fields}
and \ref{chap:DQT=000026DQC}. Alternatively, the probability amplitudes
might never be found quantized. However, instead of considering the
precision of probability amplitudes, it is better to consider the
precision of their inducing probabilities because the probability
amplitudes is unlikely to be accessed directly while the probability
of an event can be deduced after a series of experiments. Therefore,
we extend the idea of ``imprecise probabilities'' into quantum domain,
and consider quantum interval-valued probability measures (QIVPM)
in Chapter~\ref{chap:QIVPM}.

@@

Given a pure state~$\ket{\Phi}$, when we measure an observable represented
by an Hermitian matrix~$\mathbf{O}$, the measurement result is one
of the eigenvalues of $\mathbf{O}$. The probability of getting a
particular eigenvalue~$\lambda$ is $\melem{\Phi}{P}{\Phi}$, where
$P$ is the projection operator onto the eigenspace of $\lambda$.
This rule of computing the probability is called the Born rule \cite{Born1983,Mermin2007,Jaeger2007},
which is used when we want to extract information from a quantum computer.
For any mixed state~$\rho$, the generalized Born rule induces a
conventional quantum probability measure $\muB_{\rho}:\events\rightarrow[0,1]$,
where $\events$ is the set of all projection operators on a given
Hilbert space. Conversely, any quantum probability measure $\mu:\events\rightarrow[0,1]$
can be induced from a mixed state $\rho$ in the Hilbert space of
dimension $D\ge3$ according to Gleason's theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
In another word, this state $\rho$ is the unique state consistent
with any given quantum probability measure.
\end{comment}
The geometrical structure of states in CQT is first reviewed in Sec.~\ref{CQC1qubitBloch.sec},
where a pure state is represented as a vector in the Hilbert space,
and an irreducible state is actually a point in the projective Hilbert
space \cite{MosseriDandoloff2001,Jaeger2007} also known as the complex
projective space \cite{Hatcher2001,Bengtsson2007}. In Sec.~\ref{sec:fuzzy},
we will construct a quantum probability space based on classical probability
spaces, apply Gleason's theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}
to recover the Born rule \cite{Born1983,Mermin2007,Jaeger2007}, and
define the expectation value of an observable. These expectation values
will be used in Sec.~\ref{sec:The-geometry-of} to defined purity
which provides an easy way to decide whether a state is entangled
or not and to defined the maximum entangled states.

\begin{comment}
In Section~\ref{modalquantum}, we examine previously-introduced
quantum theories defined over unrestricted finite fields and show
in Section~\ref{modalquantumcomputing} that this approach leads
to theories with such bizarre powers that they are probably unphysical.
Although a version of quantum theory defined over a two-valued field
can express simple algorithms such as quantum teleportation, it is
so weak that it cannot express Deutsch's algorithm. This quantum theory
is, however, also so powerful that it can be used to solve an unstructured
database search of size $N$ using $O(\log(N))$ steps, which outperforms
the known asymptotic bound $O(\sqrt{N})$ in conventional quantum
computing.
\end{comment}
\begin{comment}
Our first discrete model replaces the complex numbers by a unrestricted
finite field~$\mathbb{F}_{q}$ \cite{Schumacher2012-SCHMQT,DQT2014,SchumacherWestmoreland2010}.
In this model, an $n$-qubit state is a non-zero vector in $\mathbb{F}_{q}^{2^{n}}$.
When we measure a given state~$\ket{\Phi}$, an observable is replaced
by a basis~$\mathcal{B}$ of $\mathbb{F}_{q}^{2^{n}}$. Since $\ket{\Phi}$
can be represented by the summation across a unique subset~$\mathcal{S}$
of the basis~$\mathcal{B}$, whether it is possible or impossible
to measure a basis vector~$\ket{i}$ depends on whether $\ket{i}$
is in $\mathcal{S}$ or not. Because this model only predicts whether
a result is possible or impossible, it is called the modal quantum
theory. Its computational model, called the modal quantum computing,
is far from conventional quantum computing. Although modal quantum
computing can express simple algorithms such as quantum teleportation
\cite{BennettBrassardEtAl1993,peres1995quantum,Mermin2007,Jaeger2007},
it is so weak that it cannot express Deutsch's algorithm \cite{Deutsch1985,Mermin2007}.
This quantum theory is, however, also so powerful that it can be used
to solve an unstructured database search of size $N$ using $O(\log(N))$
steps, which outperforms the known asymptotic bound $O(\sqrt{N})$
in conventional quantum computing \cite{Grover:1996:FQM:237814.237866,BennettBernsteinBrassardVazirani1997,Mermin2007,Jaeger2007}.
\end{comment}
In Chapter~\ref{chap:Unrestricted-Finite-Fields}, we examine previously-introduced
quantum theories defined over unrestricted finite fields \cite{Schumacher2012-SCHMQT,DQT2014,SchumacherWestmoreland2010}.
We call it the modal quantum theory because it can only predict whether
a measure result is possible or impossible, but not its probability.
Although a version of quantum theory defined over the two-valued field
can express simple algorithms such as quantum teleportation \cite{BennettBrassardEtAl1993,peres1995quantum,Mermin2007,Jaeger2007,Schumacher2012-SCHMQT},
it is so weak that it cannot express Deutsch's algorithm \cite{Deutsch1985,Mermin2007,DQT2014}.
This quantum theory is, however, also so powerful that it can be used
to solve the $\usat$ problem~\cite{Papadimitriou1993}, which is
as hard as a general $\NP$-complete problem~\cite{Valiant198685}.

\begin{comment}
Next, in Section \ref{discretequantumtheoryI}, we improve on this
by showing that for finite fields of order $p^{2}$, with the prime
$p$ of the form $4\ell+3$ ($\ell$ a non-negative integer), the
complex numbers have extremely compelling and natural discrete analogs
that permit a great many of the standard requirements of quantum computing
to be preserved. Under suitable conditions, we have amplitude-based
partitions of unity, unitary transformations, and entanglement, as
well as solutions to deterministic quantum algorithms such as the
algorithms of Deutsch, Simon, and Bernstein-Vazirani~\cite{NCbook,Mermin},
though still with some bothersome shortcomings. Because of the modular
nature of arithmetic in the finite complex field, it is not possible
to define an inner product in the usual sense, and we show in Section~\ref{discretequantumcomputingI}
that this leads to excessive computational power for the unstructured
database search problem for certain database sizes.
\end{comment}
\begin{comment}
Our second model called discrete quantum theory (I) considers only
finite fields of order $p^{2}$, with the prime $p$ of the form $4\ell+3$
($\ell$ a non-negative integer) \cite{geometry2013,DQT2014}. In
this model, an irreducible state is a vector in $\mathbb{F}_{p^{2}}^{2^{n}}$,
and can be reduced to a point in the discrete complex projective space~$\mathbb{DCP}^{2^{n}-1}$.
Among these irreducible states, the product states and entangled states
can not only be identified as in CQT but also be counted for different
$p$. Although the state space is a natural discrete analog to CQT,
this model still only predicts whether a measurement result is possible
or impossible, because it is not possible to define an inner product
in the usual sense due to the modular arithmetic~\cite{grove2002classical}.
Under suitable conditions, we can have deterministic quantum algorithms
such as the algorithms of Deutsch, Simon \cite{Simon:1994:PQC:1398518.1399019,Mermin2007,Jaeger2007},
and Bernstein-Vazirani \cite{Bernstein:1993:QCT:167088.167097,Mermin2007},
but this still leads to excessive computational power for the unstructured
database search problem for certain database sizes.
\end{comment}
In Sec.~\ref{discretequantumtheoryI} to \ref{discretequantumcomputingI},
we improve on this by showing that for finite fields of order $p^{2}$,
$\Fpp$, with the prime $p$ of the form $4\ell+3$ ($\ell$ a non-negative
integer), the complex numbers have extremely compelling and natural
discrete analogs that permit a great many of the standard requirements
of quantum computing to be preserved. In this model, the states of
a discrete $d$-dimensional system is a vector in the $d$-dimensional
vector space over $\Fpp$, and can be reduced to a point in the discrete
complex projective space. Among these irreducible states, the product
states and entangled states could not only be identified as in CQT
but also be counted for different $p$. Despite this model's similarity
to CQT, it still only predicts whether a measurement result is possible
or impossible, because it is not possible to define an inner product
in the usual sense due to the modular arithmetic~\cite{grove2002classical}.
Under suitable conditions, we can have deterministic quantum algorithms
such as the algorithms of Deutsch, Simon \cite{Simon:1994:PQC:1398518.1399019,Mermin2007,Jaeger2007},
and Bernstein-Vazirani \cite{Bernstein:1993:QCT:167088.167097,Mermin2007},
but this still leads to excessive computational power for partially
solving the $\usat$ problem.

@@

Finally, in Section~\ref{discretequantumcomputingII}, we apply our
discrete quantum theory to the study of two representative algorithms,
the deterministic Deutsch-Jozsa algorithm and the probabilistic Grover
algorithm~\cite{NCbook,Mermin}. The first algorithm highlights the
role played by the size of the field $p$ in determining the actual
resources required for computation as the number of input bits $n$
increases, a concept nonexistent in conventional quantum computing.
The second algorithm highlights, in addition, the dependence of the
precision of measurement (via cardinal probabilities) on the size
of the field, another nonexistent concept in conventional quantum
computing.
\end{comment}
\begin{comment}
Our third model, discrete quantum theory (II)~\cite{DQT2014}, restricts
states in some local region of $\mathbb{F}_{p^{2}}^{2^{n}}$. Within
a local region, a notion of inner product and probability could be
recovered. Its discrete quantum computing can be further applied to
the deterministic Deutsch-Jozsa algorithm \cite{DeutschJozsa1992,Jaeger2007}
and the probabilistic Grover algorithm \cite{Grover:1996:FQM:237814.237866,Mermin2007,Jaeger2007}.
\end{comment}
Since many important quantum algorithms are probabilistic, we want
to find a notion of probability for quantum theory over finite fields.
The cardinal probability defined in Sec.~\ref{sec:Discrete-Quantum-Theory-(II)}
restricts states in some local region of the vector spaces. Within
a local region, a notion of inner product and probability could be
recovered. In Sec.~\ref{discretequantumcomputingII}, this discrete
quantum theory is applied on two representative algorithms, the deterministic
Deutsch-Jozsa algorithm \cite{DeutschJozsa1992,Jaeger2007} and the
probabilistic Grover algorithm \cite{Grover:1996:FQM:237814.237866,Mermin2007,Jaeger2007}.
\begin{comment}
When people tried to define quantum probability over finite fields,
people tended to treat the original Born rule as an axiom, and tried
to modify it to get a discrete Born rule \cite{Schumacher2012-SCHMQT,doi:10.1142/S0217984913500644,DQT2014,Ellerman2016a}.
However, any modified Born rule could hardly work on the whole vector
space, since there is no inner product on the whole vector space over
finite fields. Instead of treating the Born rule as an axiom, the
Born rule can actually be deduced from a set of abstract definitions
and axioms according to Gleason's theorem. Although we might hope
to deduce a discrete Born rule directly from a similar set of definitions
and axioms, no discrete Born rule satisfies certain properties motivated
by Gleason's theorem with infinitely precise real-number probability~\cite{Gardiner2014}.
Since the state spaces are now discrete and finite, this suggests
us to consider a discrete Born rule mapping to finite number of intervals
called interval-valued probability~\cite{JamisonLodwick2004,THOS2017}.
To adopting the idea of interval-valued probability step-by-step,
before attempting to study quantum interval-valued probability over
finite fields, we will first review the classical interval-valued
probability, and extend it with the conventional quantum theory.
\end{comment}
Despite the great success of cardinal probabilities, it is difficulty
to define their arithmetic operations and expectation values. Since
these could be defined easily with real-valued probability measures
in CQT, we attempt to consider quantum probability measures over finite
fields (QPMFF) in Sec.~\ref{sec:Toward-IVPM}. In contrast to CQT,
our Gleason-like conditions over finite fields cannot deduce a discrete
Born rule in general. This impossible result suggests us that the
discrete and finite state spaces might not compatible with infinitely
precise real-valued probability. Thus, no matter whether the probability
amplitudes are ultimately discrete or not,

>>>>

\centering{}\centering{}%
\begin{tabularx}{\linewidth}{ccYY}
\toprule 
\addlinespace
Physical quantities & Discrete & Discrete with extreme small unit & Continuous, or no evidence to support they are discrete\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational devices & Digital computer & Analog computer & Analog computer\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Mathematical representation & Integers & Real numbers & Real numbers\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational model  & Turing machine & BCSS machine & BCSS machine\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
How the models realize? & Reliably & 
Not reliably, but after we have a better technology to manipulate the minimum units, we can predict the physical phenomenon and their computational power by a better discrete model. & 
Not reliably, because precision can never be high enough, and the difference between their computational power might not be able compensated by error analysis techniques. \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Inspired quantum models &  & Quantum theories and computing over finite fields & Quantum interval-valued probability measures (QIVPM)\tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The situations for classical quantities and the computational power
above them are summarized in Table~\ref{tab:Comparison}, and the
two possibilities dealing with real-valued models inspires us two
types of quantum models. Similar to classical analog computers, the
quantum circuit model, the most widely used model of quantum computing,
manipulates quantities which are not discrete apparently: the probability
amplitudes of quantum states. To address the similar inconsistency,
we first review the ``conventional quantum theory'' (CQT)\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.} in Chapter~\ref{chap:Conventional-Quantum-Theory}, where the probability
amplitudes are in the field of complex numbers. If the probability
amplitudes are assumed to be ultimately discrete, since their field
structure is essential to construct a quantum theory, we might assume
they are elements in discrete finite fields and build the quantum
theories and computing upon these finite fields in Chapter~\ref{chap:Unrestricted-Finite-Fields}
and \ref{chap:DQT=000026DQC}. Alternatively, the probability amplitudes
might never be found quantized so that we should try another approach
to take into account the precision of quantities. However, instead
of the precision of probability amplitudes, it is better to consider
the precision of their inducing probabilities because the probability
amplitudes is unlikely to be accessed directly while the probability
of an event can be deduced after a series of experiments. Moreover,
the idea of ``imprecise probabilities'' have well-studied classically
\cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
In Chapter~\ref{chap:QIVPM}, they will be extended it to quantum
domain, and define quantum interval-valued probability measures (QIVPM)
.

@@

Given a pure state~$\ket{\Phi}$, when we measure an observable represented
by an Hermitian matrix~$\mathbf{O}$, the measurement result is one
of the eigenvalues of $\mathbf{O}$. The probability of getting a
particular eigenvalue~$\lambda$ is $\melem{\Phi}{P}{\Phi}$, where
$P$ is the projection operator onto the eigenspace of $\lambda$.
This rule of computing the probability is called the Born rule \cite{Born1983,Mermin2007,Jaeger2007},
which is used when we want to extract information from a quantum computer.
For any mixed state~$\rho$, the generalized Born rule induces a
conventional quantum probability measure $\muB_{\rho}:\events\rightarrow[0,1]$,
where $\events$ is the set of all projection operators on a given
Hilbert space. Conversely, any quantum probability measure $\mu:\events\rightarrow[0,1]$
can be induced from a mixed state $\rho$ in the Hilbert space of
dimension $D\ge3$ according to Gleason's theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
In another word, this state $\rho$ is the unique state consistent
with any given quantum probability measure.
\end{comment}
The geometrical structure of states in CQT is first reviewed in Sec.~\ref{CQC1qubitBloch.sec},
where a pure state is represented as a vector in a Hilbert space,
and an irreducible state is actually a point in a projective Hilbert
space \cite{MosseriDandoloff2001,Jaeger2007} also known as a complex
projective space \cite{Hatcher2001,Bengtsson2007}. In Sec.~\ref{sec:fuzzy},
we will construct a quantum probability space based on classical probability
spaces, apply Gleason's theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}
to recover the Born rule \cite{Born1983,Mermin2007,Jaeger2007}, and
define the expectation value of an observable. The expectation value
will then be used in Sec.~\ref{sec:The-geometry-of} to defined purity
which provides an easy way to decide whether a state is entangled
or not and to defined the maximum entangled states.

\begin{comment}
In Section~\ref{modalquantum}, we examine previously-introduced
quantum theories defined over unrestricted finite fields and show
in Section~\ref{modalquantumcomputing} that this approach leads
to theories with such bizarre powers that they are probably unphysical.
Although a version of quantum theory defined over a two-valued field
can express simple algorithms such as quantum teleportation, it is
so weak that it cannot express Deutsch's algorithm. This quantum theory
is, however, also so powerful that it can be used to solve an unstructured
database search of size $N$ using $O(\log(N))$ steps, which outperforms
the known asymptotic bound $O(\sqrt{N})$ in conventional quantum
computing.
\end{comment}
\begin{comment}
Our first discrete model replaces the complex numbers by a unrestricted
finite field~$\mathbb{F}_{q}$ \cite{Schumacher2012-SCHMQT,DQT2014,SchumacherWestmoreland2010}.
In this model, an $n$-qubit state is a non-zero vector in $\mathbb{F}_{q}^{2^{n}}$.
When we measure a given state~$\ket{\Phi}$, an observable is replaced
by a basis~$\mathcal{B}$ of $\mathbb{F}_{q}^{2^{n}}$. Since $\ket{\Phi}$
can be represented by the summation across a unique subset~$\mathcal{S}$
of the basis~$\mathcal{B}$, whether it is possible or impossible
to measure a basis vector~$\ket{i}$ depends on whether $\ket{i}$
is in $\mathcal{S}$ or not. Because this model only predicts whether
a result is possible or impossible, it is called the modal quantum
theory. Its computational model, called the modal quantum computing,
is far from conventional quantum computing. Although modal quantum
computing can express simple algorithms such as quantum teleportation
\cite{BennettBrassardEtAl1993,peres1995quantum,Mermin2007,Jaeger2007},
it is so weak that it cannot express Deutsch's algorithm \cite{Deutsch1985,Mermin2007}.
This quantum theory is, however, also so powerful that it can be used
to solve an unstructured database search of size $N$ using $O(\log(N))$
steps, which outperforms the known asymptotic bound $O(\sqrt{N})$
in conventional quantum computing \cite{Grover:1996:FQM:237814.237866,BennettBernsteinBrassardVazirani1997,Mermin2007,Jaeger2007}.
\end{comment}
In Chapter~\ref{chap:Unrestricted-Finite-Fields}, we examine previously-introduced
quantum theories defined over unrestricted finite fields \cite{Schumacher2012-SCHMQT,DQT2014,SchumacherWestmoreland2010}.
It is called the modal quantum theory because it can only predict
whether a measure result is possible or impossible, but not its probability.
Although a version of quantum theory defined over the two-valued field
can express simple algorithms such as quantum teleportation \cite{BennettBrassardEtAl1993,peres1995quantum,Mermin2007,Jaeger2007,Schumacher2012-SCHMQT},
it is so weak that it cannot express Deutsch's algorithm \cite{Deutsch1985,Mermin2007,DQT2014}.
This quantum theory is, however, also so powerful that it can be used
to efficiently solve the $\usat$ problem~\cite{Papadimitriou1993},
which is as hard as a general $\NP$-complete problem~\cite{Valiant198685}.

\begin{comment}
Next, in Section \ref{discretequantumtheoryI}, we improve on this
by showing that for finite fields of order $p^{2}$, with the prime
$p$ of the form $4\ell+3$ ($\ell$ a non-negative integer), the
complex numbers have extremely compelling and natural discrete analogs
that permit a great many of the standard requirements of quantum computing
to be preserved. Under suitable conditions, we have amplitude-based
partitions of unity, unitary transformations, and entanglement, as
well as solutions to deterministic quantum algorithms such as the
algorithms of Deutsch, Simon, and Bernstein-Vazirani~\cite{NCbook,Mermin},
though still with some bothersome shortcomings. Because of the modular
nature of arithmetic in the finite complex field, it is not possible
to define an inner product in the usual sense, and we show in Section~\ref{discretequantumcomputingI}
that this leads to excessive computational power for the unstructured
database search problem for certain database sizes.
\end{comment}
\begin{comment}
Our second model called discrete quantum theory (I) considers only
finite fields of order $p^{2}$, with the prime $p$ of the form $4\ell+3$
($\ell$ a non-negative integer) \cite{geometry2013,DQT2014}. In
this model, an irreducible state is a vector in $\mathbb{F}_{p^{2}}^{2^{n}}$,
and can be reduced to a point in the discrete complex projective space~$\mathbb{DCP}^{2^{n}-1}$.
Among these irreducible states, the product states and entangled states
can not only be identified as in CQT but also be counted for different
$p$. Although the state space is a natural discrete analog to CQT,
this model still only predicts whether a measurement result is possible
or impossible, because it is not possible to define an inner product
in the usual sense due to the modular arithmetic~\cite{grove2002classical}.
Under suitable conditions, we can have deterministic quantum algorithms
such as the algorithms of Deutsch, Simon \cite{Simon:1994:PQC:1398518.1399019,Mermin2007,Jaeger2007},
and Bernstein-Vazirani \cite{Bernstein:1993:QCT:167088.167097,Mermin2007},
but this still leads to excessive computational power for the unstructured
database search problem for certain database sizes.
\end{comment}
In Sec.~\ref{discretequantumtheoryI} to \ref{discretequantumcomputingI},
we improve on this by showing that for finite fields of order $p^{2}$,
$\Fpp$, with the prime $p$ of the form $4\ell+3$ ($\ell$ a non-negative
integer), the complex numbers have extremely compelling and natural
discrete analogs that permit a great many of the standard requirements
of quantum computing to be preserved. In this model, the states of
a discrete $d$-dimensional system is a vector in the $d$-dimensional
vector space over $\Fpp$, and can be reduced to a point in the discrete
complex projective space. Among these irreducible states, the product
states and entangled states could not only be identified as in CQT
but also be counted for different $p$. Despite this model's similarity
to CQT, it still only predicts whether a measurement result is possible
or impossible, because it is not possible to define an inner product
in the usual sense due to the modular arithmetic~\cite{grove2002classical}.
Under suitable conditions, we can have deterministic quantum algorithms
such as the algorithms of Deutsch, Simon \cite{Simon:1994:PQC:1398518.1399019,Mermin2007,Jaeger2007},
and Bernstein-Vazirani \cite{Bernstein:1993:QCT:167088.167097,Mermin2007},
but this still leads to excessive computational power for partially
solving the $\usat$ problem efficiently.

@@

Finally, in Section~\ref{discretequantumcomputingII}, we apply our
discrete quantum theory to the study of two representative algorithms,
the deterministic Deutsch-Jozsa algorithm and the probabilistic Grover
algorithm~\cite{NCbook,Mermin}. The first algorithm highlights the
role played by the size of the field $p$ in determining the actual
resources required for computation as the number of input bits $n$
increases, a concept nonexistent in conventional quantum computing.
The second algorithm highlights, in addition, the dependence of the
precision of measurement (via cardinal probabilities) on the size
of the field, another nonexistent concept in conventional quantum
computing.
\end{comment}
\begin{comment}
Our third model, discrete quantum theory (II)~\cite{DQT2014}, restricts
states in some local region of $\mathbb{F}_{p^{2}}^{2^{n}}$. Within
a local region, a notion of inner product and probability could be
recovered. Its discrete quantum computing can be further applied to
the deterministic Deutsch-Jozsa algorithm \cite{DeutschJozsa1992,Jaeger2007}
and the probabilistic Grover algorithm \cite{Grover:1996:FQM:237814.237866,Mermin2007,Jaeger2007}.
\end{comment}
Since many important quantum algorithms are probabilistic, we want
to find a notion of probability for quantum theories over finite fields.
The cardinal probability defined in Sec.~\ref{sec:Discrete-Quantum-Theory-(II)}
restricts states in some local regions of the vector spaces. Within
a local region, a notion of inner product and probability could be
recovered. In Sec.~\ref{discretequantumcomputingII}, this discrete
quantum theory is applied on two representative algorithms, the deterministic
Deutsch-Jozsa algorithm \cite{DeutschJozsa1992,Jaeger2007} and the
probabilistic Grover algorithm \cite{Grover:1996:FQM:237814.237866,Mermin2007,Jaeger2007}.
\begin{comment}
When people tried to define quantum probability over finite fields,
people tended to treat the original Born rule as an axiom, and tried
to modify it to get a discrete Born rule \cite{Schumacher2012-SCHMQT,doi:10.1142/S0217984913500644,DQT2014,Ellerman2016a}.
However, any modified Born rule could hardly work on the whole vector
space, since there is no inner product on the whole vector space over
finite fields. Instead of treating the Born rule as an axiom, the
Born rule can actually be deduced from a set of abstract definitions
and axioms according to Gleason's theorem. Although we might hope
to deduce a discrete Born rule directly from a similar set of definitions
and axioms, no discrete Born rule satisfies certain properties motivated
by Gleason's theorem with infinitely precise real-number probability~\cite{Gardiner2014}.
Since the state spaces are now discrete and finite, this suggests
us to consider a discrete Born rule mapping to finite number of intervals
called interval-valued probability~\cite{JamisonLodwick2004,THOS2017}.
To adopting the idea of interval-valued probability step-by-step,
before attempting to study quantum interval-valued probability over
finite fields, we will first review the classical interval-valued
probability, and extend it with the conventional quantum theory.
\end{comment}
Despite the great success of cardinal probabilities, it is difficult
to define their arithmetic operations and expectation values. Since
these could be defined easily with real-valued probabilities in CQT,
we attempt to consider quantum probability measures over finite fields
(QPMFF) in Sec.~\ref{sec:Toward-IVPM}. In contrast to CQT, our Gleason-like
conditions over finite fields cannot deduce a discrete Born rule in
general. This impossible result suggests us that the discrete and
finite state spaces might not compatible with infinitely precise real-valued
probability. Thus, no matter whether the probability amplitudes are
ultimately discrete or not,
